---
title: "Diabetic Outcomes"
author: "Cormac Nolan"
date: "28 January 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(413)
```

```{r packages_and_data, warning=FALSE, message=FALSE}
library(tidyverse)
library(cluster)
library(knitr)
library(Rtsne)
library(FactoMineR)
library(factoextra)

diabetes_raw <- 
  read_csv("./data/dataset_diabetes/diabetic_data.csv", na='?') %>% 
  mutate_if(is.character, factor)
```

```{r functions}

calc_silhouette <- function(distances, k){
  silhouette_width <- c(NA)
  for(i in 2:k){
    pam_fit <- pam(distances, diss = TRUE, k = i)
    silhouette_width[i] <- pam_fit$silinfo$avg.width
  }  
  silhouette_width
}

get_gower_extremes <- function(gower_distance_object) {
  
  diabetes_gower_matrix <- gower_distance_object %>% as.matrix
  list(most_similar = 
  diabetes_clean_sample[
    which(diabetes_gower_matrix ==
            min(diabetes_gower_matrix[diabetes_gower_matrix !=
                                        min(diabetes_gower_matrix)
                                      ]
            ),
          arr.ind = TRUE)[1, ], ]
  ,
  least_similar = diabetes_clean_sample[
    which(diabetes_gower_matrix ==
            max(diabetes_gower_matrix[diabetes_gower_matrix !=
                                        max(diabetes_gower_matrix)
                                      ]
            ),
          arr.ind = TRUE)[1, ], ]
  )
}

```

## Diabetes {.tabset}
All work here based on this [r-bloggers post](https://www.r-bloggers.com/clustering-mixed-data-types-in-r/) by [Daniel P. Martin of Wicked Good Data.](https://dpmartin42.github.io/)

Data sourced from the [UCI Machine Learning Repository.](https://archive.ics.uci.edu/ml/datasets.html?area=&att=&format=&numAtt=&numIns=&sort=nameUp&task=clu&type=&view=table)


### Explore
```{r explore}
# diabetes_raw %>% View()

diabetes_raw %>% summary()

diabetes_raw %>% group_by(weight) %>% tally() # weight
diabetes_raw %>% group_by(age) %>% tally()

```
Some of the variables seem less useful for cluster, for example weight is only present for a tiny minority of cases. Various identification numbers should also be removed before clustering or modelling.


### Clean
```{r clean}
diabetes_clean <-
  diabetes_raw %>% 
  select(-payer_code, -encounter_id)

diabetes_clean_sample <- diabetes_clean %>% sample_frac(0.02)

```

### Clustering
First we calculate the distance metric. The full dataset is far too large to calculate a matrix of Gower distances for each patient so we have created a random sample of ~2k. We can then use these distances to cluster the pateints. Here we'll calculate the gower distance and then perform a quick sanity check on the most and least similar pairs.
```{r distance calculation, cache=TRUE}
diabetes_gower_dist <- 
  diabetes_clean_sample %>% select(-readmitted, -patient_nmb) %>% 
  daisy(metric="gower")

get_gower_extremes(diabetes_gower_dist) %>% lapply(kable)

```

First we need to check which cluster size works best using the Silhouette method.
```{r number of clusters}
silhouette_width <- calc_silhouette(diabetes_gower_dist, k = 10)

ggplot(data.frame(k = 1:10, silhouette_width),
       aes(x = k, y = silhouette_width)) + 
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = 2:10, labels = 2:10) +
  labs(y="Silhouette Width", title="Silhouette Width") +
  theme_bw()
```

k=4 seems to be a pretty good estimate.


```{r PAM Clustering}
pam_fit <- pam(diabetes_gower_dist, diss = TRUE, k = 4)
pam_fit_k5 <- pam(diabetes_gower_dist, diss = TRUE, k = 5)

diabetes_clean_sample %>%
  filter(row_number() %in% pam_fit$medoids) %>% kable()# View()

```

### Visualisation & Assessment
Looking at the t-SNE plot below (which allows visualisation of multi-variate data in a lower dimensional space), I have a few thoughts. There might be too many or too few clusters; there might be too many variables to adequately project into lower dimensions; there might be too many variables in general which is confusing the analysis. Next steps might be to use PCA
```{r}
diabetes_tsne <- Rtsne(diabetes_gower_dist, is_distance = TRUE)
diabetes_tsne_k5 <- Rtsne(diabetes_gower_dist, is_distance = TRUE)

diabetes_tsne_data <- diabetes_tsne$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit$clustering),
         patient_nbr = diabetes_clean_sample$patient_nbr) 

diabetes_tsne_data_k5 <- diabetes_tsne_k5$Y %>%
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit_k5$clustering),
         patient_nbr = diabetes_clean_sample$patient_nbr) 

diabetes_tsne_data %>% 
  ggplot(aes(x = X, y = Y)) +
  geom_point(aes(color = cluster)) +
  theme_bw() +
  labs(title="t-SNE Plot for Diabetic Clustering - k = 4")

diabetes_tsne_data_k5 %>% 
  ggplot(aes(x = X, y = Y)) +
  geom_point(aes(color = cluster)) +
  theme_bw() +
  labs(title="t-SNE Plot for Diabetic Clustering - k = 5")

diabetes_tsne_data %>% 
  filter(X > 20 & X < 26,
         Y > 5 & Y < 15) %>%
  left_join(diabetes_clean_sample, by = "patient_nbr")

```

## Multiple Correspondence Analysis
```{r MCA}
diabetes_sample_mfa <-
  diabetes_clean_sample %>% select(-weight)
  MFA(group = c(1, 4, 3, 1, 
                1, 6, 3,
                1, 27, 1),
      type = c("n", "n", "n", "s", 
               "n", "s", "n",
               "s", "n", "n"),
      name.group=c("patient", "demographics", "admission", "admission_dur",
                   "specialty", "procedures", "diagnoses",
                   "num_diagnoses", "medication", "outcome"),
      num.group.sup = c(1, 10),
      graph = FALSE)



```